A Switching Planner for Combined Task and Observation Planning

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
tiny changes.

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
language that facilitates factored description of domains and problem
instances. There are straightforward compilations from problems
expressed compactly in PPDDL to propositional representations amenable
to state-of-the-art planning procedures.

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called DTPDDL. This can
express probabilistic models of the sensing consequences of acting, to
quantitatively capture unreliability in perception. That expressive
power is achieved by incorporating perceptual analogues of fluent,
predicate, and action definitions. In detail, we have declarations of
state characterising predicate and fluent symbols according to the
PPDDL syntax. In addition, we allow two other declarations, for
perceptual predicates and fluents respectively. For example, suppose
our robot is tasked with exploring {\em locations} in order to
identify the whereabouts of a {\em visual-object}. We must describe
state and perceptual facts that model the {\em true}, resp. perceived,
locations of objects. In DTPDDL, these declarations appear as:

@@

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect on the state. That is, its
execution incurs an instantaneous penalty, $3$, that corresponds to
the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
reward} is a {\em reserved word}, and occurs in {\em increased} and
{\em decreased} terms in operator schemata. In that setting, {\em
reward} is interpreted as accumulated (resp. instantaneous) rewards
and penalties.}

@@

\noindent Here, the {\em execution} declaration is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid us when discussing
our switching continual planner. The distribution is expressed in a
tree-like structure of terms. Each term is either: (1) atomic, e.g., a
state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
atomic terms. The root term is always conjunctive, and the leaves are
atomic. For example, the starting distribution for $\pp{DORA}$ might
be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
expresses the conjunctive root of the tree -- I.e., the root node
$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
subterm.}

>>>>

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
language that facilitates factored descriptions of domains and problem
instances. There are straightforward compilations from problems
expressed compactly in PPDDL to propositional representations amenable
to state-of-the-art planning procedures.

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called decision-theoretic
(DT)PDDL. This can express probabilistic models of the sensing
consequences of acting, to quantitatively capture unreliability in
perception. That expressive power is achieved by incorporating
perceptual analogues of fluent, predicate, and action definitions. In
detail, we have declarations of state characterising predicate and
fluent symbols according to the PPDDL syntax. In addition, we allow
two other declarations, for perceptual predicates and fluents
respectively. For example, suppose our robot is tasked with exploring
{\em locations} in order to identify the whereabouts of a {\em
visual-object}. We must describe state and perceptual facts that model
the {\em true}, resp. perceived, locations of objects. In DTPDDL,
these declarations appear as:

@@

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect on the state. That is, its
execution incurs an instantaneous penalty, $3$, that corresponds to
the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
reward} is a {\em reserved word}, and occurs in {\em increased} and
{\em decreased} terms in operator schemata. In that setting, {\em
reward} is interpreted as accumulated, whereas in DTPDDL it is
instantaneous.}

@@

\noindent Here, the {\em execution} declaration is a lifted description of
the sense action-precondition -- i.e,
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid us when discussing
our switching continual planner. The distribution is expressed in a
tree-like structure of terms. Each term is either: (1) atomic, e.g., a
state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
atomic terms. The root term is always conjunctive, and the leaves are
atomic. For example, the starting distribution for $\pp{DORA}$ might
be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
expresses the conjunctive root of the tree -- i.e., the root node
$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
subterm.}

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
tiny changes.

We now describe a continual planning approach that switches between
sequential and decision-theoretic planning sessions.

@@

In a deterministic-action POMDP, all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
do define a set of state-assumptions available to sequential planning.
%%
Writing \#\ if the value of a proposition is unspecified in an
assumption, for our $\pp{DORA}$ example we have that the following
assumptions:

@@

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.g., for
$\pp{DORA}$:

@@

\noindent which is equal to the maximal contribution a trace of the plan can
make to Eq~\ref{eq:expectedvalue}. %% In other words, writing $V_i$ for
%% the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
%% s_1, a_1,.., s_N$, we have
%%
%% \[
%% V^* = \arg\min_{V_i} \Expect \bigg{[} 
%% \sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
%% \bigg{]} - V_i.
%% \]
%%
In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly jettisoning a
false assumption. The added dual actions are propositionally trivial
-- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic session, we include another propositionally trivial
action $\actions.\poss(\action_i)$ with the reward property:

@@

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. Each iteration, we add one candidate
term with the largest number of parent terms already in the candidate
tree.

>>>>

We now describe our continual planning approach that switches between
sequential and decision-theoretic planning sessions.

@@

In a deterministic-action POMDP, all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
to define a set of state-assumptions available to sequential planning.
%%
Writing \#\ if the value of a proposition is unspecified in an
assumption, for our $\pp{DORA}$ example we have the following
assumptions:

@@

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- e.g., for
$\pp{DORA}$:

@@

\noindent which is equal to the maximal contribution a trace of the plan can
make to Eq~\ref{eq:expectedvalue}. %% In other words, writing $V_i$ for
%% the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
%% s_1, a_1,.., s_N$, we have
%%
%% \[
%% V^* = \arg\min_{V_i} \Expect \bigg{[} 
%% \sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
%% \bigg{]} - V_i.
%% \]
%%
In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of
\system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly jettisoning a
false assumption. The added dual actions are propositionally trivial
-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic session, we include another propositionally trivial
action $\actions.\poss(\action_i)$ with the reward property:

@@

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. In each iteration, we add one
candidate term with the largest number of parent terms already in the
candidate tree.
